### 2016-05-12 text/Chapter08.tex
Edits in implementation, consistency (Chapter 7)

\noindent
A consequence of the equivalence discussed above is that the unified coeffect system can capture
all properties that can be captured by the flat coeffect system -- including implicit parameters,
rebindable resources, Haskell type classes (discussed by Orchard~\cite{comonads-dom-thesis}),
data-flow and variable liveness.

@@

\item Another approach that works for most (but not all) systems we discussed is to require the structure
$(\C,\cpar,\czip)$ to form a lattice. This is consistent with all our structural coeffect systems as
well as flat systems for liveness and data-flow, but not with the system for tracking of implicit
parameters (or sets of rebindable resources).
\end{itemize}

>>>>

\noindent
A consequence of the equivalence discussed above is that the unified coeffect system can capture
all properties that can be captured by the flat coeffect system -- including implicit parameters,
rebindable resources, Haskell type classes (discussed by Orchard~\cite{comonads-dom-thesis}),
dataflow and variable liveness.

@@

\item Another approach that works for most (but not all) systems we discussed is to require the structure
$(\C,\cpar,\czip)$ to form a lattice. This is consistent with all our structural coeffect systems as
well as flat systems for liveness and dataflow, but not with the system for tracking of implicit
parameters (or sets of rebindable resources).
\end{itemize}

### 2016-05-12 text/Chapter09.tex
Fix references
Edits in implementation, consistency (Chapter 7)

This thesis aims to change how developers and programming language designers think about
\emph{context} or \emph{execution environments} of programs. Such context or execution environment
has numerous forms -- mobile applications access network, GPS location or user's personal data;
journalists obtain information published on the web or through open government data initiatives;
in data-flow programming, we have access to past values of expressions.
This thesis aims to change our understanding of \emph{context} so that the above examples are
viewed uniformly through a single programming language abstraction, which we call \emph{coeffects},
rather than as disjoint cases.

@@

In $\lambda$-calculus, the term \emph{context} is usually refers to the free-variable context.
However, other programming language features are also related to context or program's
execution environment. In Chapter~\ref{ch:applications}, we revisit many of such features --
including resource tracking in distributed computing, cross-platform development, data-flow
programming and liveness analysis, but also Haskell's implicit parameters.

@@

In Chapter~\ref{ch:flat}, we develop a \emph{flat coeffect calculus} which gives us an unified way
of tracking \emph{whole-context} properties. The calculus is parameterized by a \emph{flat coeffect
algebra} that captures the algebraic properties of contextual information. Concrete
instances of flat coeffects include Haskell's implicit parameters, whole-context liveness and
whole-context data-flow.

@@

We unify systems with \emph{per-variable} contextual properties in Chapter~\ref{ch:structural},
which describes our \emph{structural coeffect calculus} (Section~\ref{sec:struct-calculus}).
Similarly to the flat variant, the calculus is parameterized by a \emph{structural coeffect algebra}.
Concrete instances of the calculus track bounded variable use (\ie~how many times is a variable
accessed), data-flow properties (how many past values are needed) and liveness (\ie~can variable
be accessed).

@@

The Chapters~\ref{ch:flat} and \ref{ch:structural} present the main novel technical contributions of this
thesis. In Chapter~\ref{ch:unified}, we discuss the similarities and differences between the two.
Although the calculi are worth considering separately (as they have different equational properties),
they can be presented in a uniform way. In Section~\ref{sec:unified-unified}, we present the
\emph{unified coeffect calculus}, which can be instantiate to track both per-variable and
whole-context properties. The key trick is to parameterize the calculus by the \emph{shape} of
a context, which can be a vector (corresponding to the vector of free variables) or a singleton
container (attaching one annotation to the whole context).

@@

This thesis aims to provide a similar unifying theory and tools for capturing context-dependence
in programming languages. We showed that programming language features (liveness, data-flow,
implicit parameters, etc.) that were previously treated separately can be captured by a common
framework developed in this thesis. The main technical contribution of this thesis is that it
provides the necessary tools for programming language designers -- including parameterized type
systems, categorical semantics based on indexed comonads and equational theory.

>>>>

This thesis aims to change how developers and programming language designers think about
\emph{context} or \emph{execution environments} of programs. Such context or execution environment
has numerous forms -- mobile applications access network, GPS location or user's personal data;
journalists obtain information published on the web or through open government data initiatives;
in dataflow programming, we have access to past values of expressions.
This thesis aims to change our understanding of \emph{context} so that the above examples are
viewed uniformly through a single programming language abstraction, which we call \emph{coeffects},
rather than as disjoint cases.

@@

In $\lambda$-calculus, the term \emph{context} is usually refers to the free-variable context.
However, other programming language features are also related to context or program's
execution environment. In Chapter~\ref{ch:applications}, we revisit many of such features --
including resource tracking in distributed computing, cross-platform development, dataflow
programming and liveness analysis, but also Haskell's implicit parameters.

@@

In Chapter~\ref{ch:flat}, we develop a \emph{flat coeffect calculus} which gives us an unified way
of tracking \emph{whole-context} properties. The calculus is parameterized by a \emph{flat coeffect
algebra} that captures the algebraic properties of contextual information. Concrete
instances of flat coeffects include Haskell's implicit parameters, whole-context liveness and
whole-context dataflow.

@@

We unify systems with \emph{per-variable} contextual properties in Chapter~\ref{ch:structural},
which describes our \emph{structural coeffect calculus} (Section~\ref{sec:struct-calculus}).
Similarly to the flat variant, the calculus is parameterized by a \emph{structural coeffect algebra}.
Concrete instances of the calculus track bounded variable use (\ie~how many times is a variable
accessed), dataflow properties (how many past values are needed) and liveness (\ie~can variable
be accessed).

@@

The Chapters~\ref{ch:flat} and \ref{ch:structural} present the main novel technical contributions of this
thesis. In Chapter~\ref{ch:further}, we discuss the similarities and differences between the two.
Although the calculi are worth considering separately (as they have different equational properties),
they can be presented in a uniform way. In Section~\ref{sec:further-unified}, we present the
\emph{unified coeffect calculus}, which can be instantiate to track both per-variable and
whole-context properties. The key trick is to parameterize the calculus by the \emph{shape} of
a context, which can be a vector (corresponding to the vector of free variables) or a singleton
container (attaching one annotation to the whole context).

@@

This thesis aims to provide a similar unifying theory and tools for capturing context-dependence
in programming languages. We showed that programming language features (liveness, dataflow,
implicit parameters, etc.) that were previously treated separately can be captured by a common
framework developed in this thesis. The main technical contribution of this thesis is that it
provides the necessary tools for programming language designers -- including parameterized type
systems, categorical semantics based on indexed comonads and equational theory.

### 2016-05-12 text/backup/Chapter01.tex
Edits in implementation, consistency (Chapter 7)

\todo{We present three calculi that model common notions of context-dependence and can be used
as basis for developing context-aware programming langauges with static type systems.}

>>>>

\todo{We present three calculi that model common notions of context-dependence and can be used
as basis for developing context-aware programming languages with static type systems.}
